---
title: "PM2.5 Prediction in China cities"
author: "Yu Ding, Sirui Jiang, Aileen Kate Fu, Kangning Chen"
date: "April 16, 2020"
output:
  word_document: default
  pdf_document: default
  html_document:
    theme: cosmo
    highlight: monochrome
    toc: ture
    toc_float: false
    toc_depth: 4
    number_sections: true
    code_folding: hide
---

## Part 1: Introduction

background domain knowledge of your team and importance of the problem

questions that focused and guided you; any initial hunches or hypotheses

```{r,echo = FALSE,results='hide',warning=FALSE,message=FALSE}
# load packages
library(GGally)
library(ggplot2)
library(car)
library(leaps)
library(stargazer)
library(dplyr)
library(psych)
library(forecast)
library(caTools)
library(mlbench)
library(caret)
library(ROCR)
library(class)
library(egg)
library(rpart)
library(rattle)
library(randomForest)
library(rpart.plot)
library(RColorBrewer)
library(adabag)
library(FNN)
library(gains)
library(clValid)
library(cluster)
library(factoextra)
library(epiDisplay)
library(dendextend)
```

## Part 2: Data Description

data sources (provenance, lineage); how well documented?

data preparation (imputation, normalization, transformation), data quality assessment / wrangling

target variable(s) chosen and justification

descriptive statistics and data visualizations

```{r,echo = FALSE,results='hide',warning=FALSE,message=FALSE}
beijing <- read.csv("beijingPM2.5.csv")
beijing$air <- factor(ifelse(beijing$PM_US.Post<50,"Good","Bad"))
beijing$PM2.5 <- beijing$PM_US.Post
beijing$seasonadj <- factor(beijing$season,levels = c(1,2,3,4),labels = c("Spring","Summer","Fall","Winter"))
beijing$SW <- ifelse(substr(beijing$cbwd, 1, 2)=="SW", 1, 0)
beijing$SE <- ifelse(substr(beijing$cbwd, 1, 2)=="SE", 1, 0)
beijing$NW <- ifelse(substr(beijing$cbwd, 1, 2)=="NW", 1, 0)
beijing$NE <- ifelse(substr(beijing$cbwd, 1, 2)=="NE", 1, 0)
beijing$log <- log(beijing$PM_US.Post)
options(scipen=999)
# write.csv(beijing,"C:/Users/dingy/Desktop/rcode/datarobot.csv")
```

```{r,echo = FALSE,results='hide',warning=FALSE,message=FALSE}
# Data Description
# Categorical variables
categ <- names(beijing) %in% c("seasonadj","SW","SE","NE","NW","cbwd",'air')
subdata <- beijing[!categ] # only numerical data
# stargazer(subdata, type = 'text')

# Compute median for all numeric variables
data.frame(median=sapply(subdata, median))

# Create a function to calculate mode
getmode <- function(v){
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}
# Compute mode for all numeric variables
pm2.5mod <- getmode(beijing$PM_US.Post)
yearmod <- getmode(beijing$year)
monthmod <- getmode(beijing$month)
daymod <- getmode(beijing$day)
hourmod <- getmode(beijing$hour)
seasonmod<- getmode(beijing$season)
dewpmod <- getmode(beijing$DEWP)
humimod <- getmode(beijing$HUMI)
presmod <- getmode(beijing$PRES)
tempmod <- getmode(beijing$TEMP)
iwsmod <- getmode(beijing$Iws)
precmod <- getmode(beijing$precipitation)
airmod<-getmode(beijing$air)
```

### 2.1 Boxplots for Categorical Variables

```{r,echo = FALSE,results='hide',warning=FALSE,message=FALSE}
ggplot(beijing,aes(x=PM_US.Post))+
  geom_histogram()+
  xlim(0,450)+
  labs(
    y = "Number",
    title = "Histogram of PM 2.5 in Beijing",
    x = "PM 2.5"
  )+
  theme(
    panel.background = element_rect(fill = "snow2"),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15),
    panel.grid = element_line(color="snow",size=0.5,linetype = "dashed")
  ) 
# The plot is skewed to the right. Therefore, use log(PM2.5) when building models might be a better option.

## Plot the relationship between target variable and categorical variables
ggplot(beijing,aes(x=cbwd,y=PM_US.Post))+
  geom_boxplot()+
  labs(
    x="Combined Wind Direction",
    y="PM 2.5",
    title="Boxplot for wind direction in Beijing"
  )+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

## A closer look
ggplot(beijing,aes(x=cbwd,y=PM_US.Post))+
  geom_boxplot()+
  labs(
    x="Combined Wind Direction",
    y="PM 2.5",
    title="Boxplot for wind direction in Beijing"
  )+
  ylim(0,120)+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )
# The wind directions from south have a higher average than north winds have.

ggplot(beijing,aes(x=seasonadj,y=PM_US.Post))+
  geom_boxplot()+
  labs(
    x="Season",
    y="PM 2.5",
    title="Boxplot for season in Beijing"
  )+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

ggplot(beijing,aes(x=seasonadj,y=PM_US.Post))+
  geom_boxplot()+
  labs(
    x="Season",
    y="PM 2.5",
    title="Boxplot for season in Beijing"
  )+
  ylim(0,120)+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

ggplot(beijing,aes(x=air,y=PM_US.Post))+
  geom_boxplot()+
  labs(
    x="Good/Bad",
    y="PM 2.5",
    title="Boxplot for Good/Bad air condition in Beijing with PM2.5"
  )+
  ylim(0,120)+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

ggplot(beijing,aes(x=air,y=DEWP))+
  geom_boxplot()+
  labs(
    x="Good/Bad",
    y="Dew Point",
    title="Boxplot for Good/Bad air condition in Beijing with dew point"
  )+
  ylim(0,35)+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

ggplot(beijing,aes(x=air,y=HUMI))+
  geom_boxplot()+
  labs(
    x="Good/Bad",
    y="Humidity",
    title="Boxplot for Good/Bad air condition in Beijing with humidity"
  )+
  ylim(0,100)+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

ggplot(beijing,aes(x=air,y=PRES))+
  geom_boxplot()+
  labs(
    x="Good/Bad",
    y="Air Pressure",
    title="Boxplot for Good/Bad air condition in Beijing with Air Pressure"
  )+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

ggplot(beijing,aes(x=air,y=TEMP))+
  geom_boxplot()+
  labs(
    x="Good/Bad",
    y="Temperature",
    title="Boxplot for Good/Bad air condition in Beijing with Temperature"
  )+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

ggplot(beijing,aes(x=air,y=Iws))+
  geom_boxplot()+
  labs(
    x="Good/Bad",
    y="Wind Speed",
    title="Boxplot for Good/Bad air condition in Beijing with Wind Speed"
  )+
  ylim(0,10)+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

tabpct(beijing$cbwd,beijing$air,xlab="Wind Direction",ylab="Air Condition",main = "Distribution of Air Condition by Wind Direction")
tabpct(beijing$seasonadj,beijing$air,xlab="Season",ylab="Air Condition",main = "Distribution of Air Condition by Season")
```

### 2.2 Scatter Plots for Numeric Variables

```{r,echo = FALSE,results='hide',warning=FALSE,message=FALSE}
list <- beijing[,c("PM2.5","HUMI","PRES","TEMP","Iws","precipitation","DEWP")]
# Correlation between variables
ggcorr(list, method = c("everything", "pearson"),low = "blueviolet", mid = "white", high = "darkgoldenrod3")

s1 <- ggplot(beijing, aes(y= PM2.5, x=DEWP))+
  geom_point(size = 0.1,color="rosybrown",alpha=0.8)+geom_smooth(color ='seagreen',size = 1)+ 
  theme(
    panel.background = element_rect(fill = "snow"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

s2 <- ggplot(beijing, aes(y= PM2.5, x=HUMI))+
  geom_point(size = 0.1,color="rosybrown",alpha=0.8)+geom_smooth(color ='seagreen',size = 1)+ 
  theme(
    panel.background = element_rect(fill = "snow"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

s3 <- ggplot(beijing, aes(y= PM2.5, x=PRES))+
  geom_point(size = 0.1,color="rosybrown",alpha=0.8)+geom_smooth(color ='seagreen',size = 1)+ 
  theme(
    panel.background = element_rect(fill = "snow"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

s4 <- ggplot(beijing, aes(y= PM2.5, x=TEMP))+
  geom_point(size = 0.1,color="rosybrown",alpha=0.8)+geom_smooth(color ='seagreen',size = 1)+ 
  theme(
    panel.background = element_rect(fill = "snow"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

s5 <- ggplot(beijing, aes(y= PM2.5, x=Iws))+
  geom_point(size = 0.1,color="rosybrown",alpha=0.8)+geom_smooth(color ='seagreen',size = 1)+ 
  theme(
    panel.background = element_rect(fill = "snow"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

s6 <- ggplot(beijing, aes(y= PM2.5, x=precipitation))+
  geom_point(size = 0.1,color="rosybrown",alpha=0.8)+geom_smooth(color ='seagreen',size = 1)+ 
  theme(
    panel.background = element_rect(fill = "snow"),
    panel.grid = element_blank(),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )

# Scatter plots between variables
ggarrange(s1,s2,s3,s4,s5,s6,ncol=2,nrow=3)
```

## Part 3. Multiple Regression Models

comparison and contrast between Hand-Crafted Models and DataRobot Models

management of outliers / overfitting / target leakage

interpretation of results (coefficients, rules, trees, etc.)

### 3.1 Regression Process

### 3.2 Regression Interpretation

### 3.3 Comparison with DataRobot

```{r,echo = FALSE,results='hide',warning=FALSE,message=FALSE}
# Train valid split
set.seed(66)
train.index<- sample(c(1:nrow(beijing)),nrow(beijing)*0.8)
train <- beijing[train.index,]
valid <- beijing[-train.index,]

# Delete outliers
outlier.index <- c(which(train.index==13196),which(train.index==20833),which(train.index==12939),which(train.index==13037),which(train.index==5936),which(train.index==22344),which(train.index==20832),which(train.index==12859),which(train.index==22244),which(train.index==47897),which(train.index==47896),which(train.index==39045),which(train.index==46506),which(train.index==37045),which(train.index==28896),which(train.index==38692),which(train.index==47932),which(train.index==20573),which(train.index==13197),which(train.index==20174),which(train.index==16579),which(train.index==13159),which(train.index==46448),which(train.index==12168),which(train.index==37077),which(train.index==19698))

train.df <- train[-outlier.index,]

# Best model
model.1 <- lm(log(PM_US.Post) ~ HUMI + seasonadj + Iws + cbwd + precipitation + I(PRES^2) + PRES + I(Iws*SW) + I(Iws*SE) + I(Iws*NW),data = train.df)
summary(model.1)
AIC(model.1)
par(mfrow=c(1,1))
plot(model.1)

# Regression accuracy
validation <- predict(model.1,valid)
accuracy(model.1$fitted.values,train.df$log)
accuracy(validation,valid$log)

# Histogram of regression residuals
ggplot(model.1,aes(x=model.1$residuals))+
  geom_histogram(aes(y=..density..),bins = 30,colour="grey25",fill="white")+
  stat_function(fun = dnorm,alpha=0.8,lwd=0.8, args = list(mean = mean(model.1$residuals), sd = sd(model.1$residuals)))+
    geom_density(alpha=0.2,lwd=0.8,color="salmon")+
  labs(
    x= "Residuals",
    y="",
    title = "Histogram of Regression Residuals"
  )+
  theme(
    panel.background = element_rect(fill = "snow2"),
    panel.grid = element_line(color="snow",size=0.5,linetype = "dashed"),
    axis.text.x = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.text.y = element_text(family = "Times",face = "bold",size = 10,color = "#993333"),
    axis.line = element_line(color = "#993333",size = 1,linetype = "solid"),
    text = element_text(color = "thistle4",face = "bold"),
    plot.title = element_text(hjust = 0.5,size=15)
  )
```

## Part 4. Classification Models

### 4.1 Logistic Regression

```{r,echo = FALSE,results='hide',warning=FALSE,message=FALSE}
# train valid split
set.seed(66)
train.index<- sample(c(1:nrow(beijing)),nrow(beijing)*0.7)
train <- beijing[train.index,]
valid <- beijing[-train.index,]

outlier.index <- c(which(train.index==13196),which(train.index==32319),which(train.index==32322),which(train.index==22244),which(train.index==46506),which(train.index==37045),which(train.index==32321),which(train.index==32320),which(train.index==20832),which(train.index==12939),which(train.index==28896),which(train.index==13159),which(train.index==32247),which(train.index==12859),which(train.index==37077),which(train.index==20573),which(train.index==13197),which(train.index==12168),which(train.index==20834),which(train.index==38903))

train.df <- train[-outlier.index,]

model.2 <- glm(air ~ HUMI + PRES + I(PRES^2) + cbwd + Iws + precipitation + I(precipitation^2) + seasonadj , family=binomial(link="logit"),train.df)
summary(model.2)
plot(model.2)

p <- predict(model.2,valid,type="response")
p.train <- predict(model.2,train.df,type = "response")
GorB <- ifelse(p>0.56,"Good","Bad")
GorB.train <- ifelse(p.train>0.56,"Good","Bad")
p_class <- factor(GorB,levels=levels(valid$air))
p_class.train <- factor(GorB.train,levels=levels(train.df$air))
confusionMatrix(p_class.train,train.df$air,positive = "Bad")
confusionMatrix(p_class,valid$air,positive = "Bad")
colAUC(p,valid$air,plotROC=TRUE)
```

### 4.2 Decision Tree

```{r,echo = FALSE,results='hide',warning=FALSE,message=FALSE}
set.seed(66)
train.index<- sample(c(1:nrow(beijing)),nrow(beijing)*0.7)
train <- beijing[train.index,]
valid <- beijing[-train.index,]

set.seed(1)
tree <- rpart(air ~ DEWP + HUMI + PRES + I(PRES^2) + TEMP + cbwd + Iws + precipitation + I(precipitation^2) + seasonadj, data = train, method = "class", cp = 0.00001, minsplit = 1, xval = 5)
length(tree$frame$var[tree$frame$var == "<leaf>"])
# printcp(tree.2) # Print out the cp table of cross-validation errors.

pruned.tree <- prune(tree, cp = 0.000559203 )
fancyRpartPlot(pruned.tree)

## Check Confusion Matrix for training set
pred.pruned.tree <- predict(pruned.tree, train, type='class') 
confusionMatrix(pred.pruned.tree, train$air,positive = "Bad")
## Repeat the process for test set
pred.pruned.tree.valid <- predict(pruned.tree, valid, type='class')
confusionMatrix(pred.pruned.tree.valid, valid$air,positive = "Bad")

# plot roc curve
prob <- predict(pruned.tree,valid,type = "prob")[,2]
pred <- prediction(prob,valid$air)
perf1 <- performance(pred,"tpr","fpr")
plot(perf1)

perf2 <- performance(pred,"auc")
print(perf2@y.values[[1]])
```

## Part 5. Conclusions and Reflections

overall conclusions, insights, surprises

reflections about the journey (process, roadblocks; pivoting, iterating)

professionalism of writing / visualizations / organization / delivery
